\protect Connecting every bit of knowledge: \\ The structure of Wikipedia's Flrst Link Network

### 2016-04-22 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
universal template and bibtex format changed

\documentclass[pre,twocolumn,twoside,superscriptaddress,floatfix, aps, 10pt]{revtex4-1}

\usepackage{currfile}
\lefthyphenmin=3
\righthyphenmin=2

\usepackage{graphicx,epsfig,verbatim,enumerate}
\usepackage{amssymb,amsmath}
\usepackage{ifthen}

% Fonts and typesetting settings
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\linespread{1.05} % Palatino needs more space between lines
\usepackage{microtype}
\usepackage{textcomp}

\usepackage{longtable}

\usepackage{mathtools}

\newboolean{twocolswitch}

\newcommand{\www}[1]{\url{#1}}
\newcommand{\req}[1]{(\ref{#1})}
\newcommand{\Req}[1]{Eq.~(\ref{#1})}

% Lettrines
\usepackage{lettrine}

%useful shortcuts
\def\R{\ensuremath{\mathbb{R}}} %\ensuremath adds math mode, if forgotten
\def\Q{\ensuremath{\mathbb{Q}}}
\def\N{\ensuremath{\mathbb{N}}}
\def\Z{\ensuremath{\mathbb{Z}}}
\def\C{\ensuremath{\mathbb{C}}}

%shorcuts with arguments
\newcommand{\abs}[1]{\left\vert#1\right\vert} %nice absolute values
\newcommand{\bt}[1]{\textbf{#1}} %bold
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations
\newcommand{\norm}[1]{\left\lVert#1\right\rVert} %vector norm
\renewcommand{\eq}[1]{\begin{align*}#1\end{align*}} %aligned equations

%piecewise function

%\begin{displaymath}
%   f(x) = \left\{
%     \begin{array}{lr}
%       1 & : x \in \mathbb{Q}\\
%       0 & : x \notin \mathbb{Q}
%     \end{array}
%   \right.
%\end{displaymath}

%environment
\newcommand{\tab}{\phantom{ssss}}

\usepackage{hyperref}
\usepackage{color}
\newcommand{\todo}[1]{\noindent\textcolor{blue}{{$\Box$ #1}}}

%colors
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} %dark green color
\definecolor{lightblue}{rgb}{0.149,0.545,0.824} %solarized blue
\definecolor{sred}{rgb}{0.863, 0.196, 0.184} %solarized red

\newcommand{\blue}[1]{{\leavevmode\color{lightblue}{#1}}} %solarized blue 
\newcommand{\green}[1]{{\leavevmode\color{javagreen}{#1}}} %command for green
\newcommand{\red}[1]{{\leavevmode\color{sred}{#1}}} %solarized red
\newcommand{\gray}[1]{{\leavevmode\color[gray]{0.5}{#1}}} %gray text
\newcommand{\Prob}[1]{{\rm Pr}\left(#1\right)}

\setboolean{twocolswitch}{true}

\begin{document}

\title{\protect
Connecting every bit of knowledge: \\
The structure of Wikipedia's Flrst Link Network
}

\date{\today}

\author{
\firstname{Mark}
\surname{Ibrahim}
}
\email{mark.s.ibrahim@uvm.edu}

\affiliation{Department of Mathematics \& Statistics, 
    Computational Story Lab,  \\
    Vermont Complex Systems Center,
    Vermont Advanced Computing Core,
    The University of Vermont, Burlington, VT 05401.}

\author{
\firstname{Christopher}
\surname{M. Danforth}
}
\email{chris.danforth@uvm.edu}

\affiliation{Department of Mathematics \& Statistics, 
    Computational Story Lab,  \\
    Vermont Complex Systems Center,
    Vermont Advanced Computing Core,
    The University of Vermont, Burlington, VT 05401.}

\author{
\firstname{Peter}
\surname{Sheridan Dodds}
}

\email{peter.dodds@uvm.edu}

\affiliation{Department of Mathematics \& Statistics, 
    Computational Story Lab,  \\
    Vermont Complex Systems Center,
    Vermont Advanced Computing Core,
    The University of Vermont, Burlington, VT 05401.}

\begin{abstract}
  \protect
Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
By following the first link in each article, we algorithmically construct a directed network of 
all 4.7 million articles: Wikipedia's First Link Network.
Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
articles on inventions, places, people, objects, and events are related and organized.  
We traverse every path, measuring the accumulation of first links, path lengths, basins, cycles, 
and the influence each article exerts in shaping the network.
We find scale-free distributions describe path length, 
accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
at a few articles---flowing from specific to general and culminating around fundamental notions such as
Community, State, and Science. 
Philosophy funnels more paths than any other article by two orders of magnitude. 
We also observe a gravitation towards topical articles such as 
Health Care and Fossil Fuel. 
These findings enrich our view of the connections and structure of
Wikipedia's ever growing store of knowledge.

\end{abstract}

\maketitle

%TODO:
% Abstract 
    % In traversing every path, we measure ?
    % Our findings... braided network (what does this mean?)
% Formatting
    % incorporate into template
% Intro
    % add month to monthly view stats
% Method
    % what to cite for drainage basin in geomorphology

% continue on Page 3

@@

Wikipedia is a towering achievement of the Internet age. 
At no point in history has a larger or more meticulously indexed collection of human knowledge 
existed.
Wikipedia contains 37 million articles in 283 languages, 
with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
\cite{drugs, stats}.
Demonstrating its relevance to modern inquiry,
Wikipedia is the sixth most visited
site in the world, surpassing 18 billion page views and 11 million edits in a single month.
\cite{wiki_edits, wiki_views}.

Wikipedia has naturally become the object of many studies. 
Researchers have examined the cultural dynamics among editors
\cite{editors},
the accuracy of the content relative to traditional encyclopedias
\cite{accuracy1, accuracy2},
the topics covered 
\cite{coverage},
and bias against portions of the population
\cite{bias_women}.
Wikipedia's content has also proven to be a powerful tool. 
Researchers have used Wikipedia to identify missing dictionary entries
\cite{missing_entries},
cluster short text
\cite{clustering},
compute semantic relatedness
\cite{semantic_relatedness},
and disambiguate meaning \cite{disambiguating}.

While these many studies have dissected and fruitfully applied Wikipedia's content,
here we examine the connections among the many articles.
A hyperlink from one Wikipedia article to another
naturally indicates a relationship between the two articles
\cite{relevance}.
The notion that hyperlinks convey information about the content of a
page has proved enormously successful in multiple domains from search engine algorithms 
such as PageRank 
\cite{pagerank} 
to topic classification
\cite{classifier}.
Here, we treat a hyperlink as a mechanism connecting two topics.

@@

By following the first hyperlink contained in all articles of the English edition of
Wikipedia as of November, 2014, we form a directed network:
{\it Wikipedia's First Link Network} (FLN).
Inspired by the claim that the majority of first links lead to
``Philosophy''---popularized by an 
\href{https://xkcd.com/903/}{xkcd}
comic and subsequently
discussed in blog posts 
\cite{xkcd, mat_blog, Ilmari_first_links}
---
we holistically study how the FLN  yields a flow of connections.
For methodological details, see Sec.~\ref{Appendix}.

The FLN is a wealth of relations among inventions, places,
figures, objects, and events across space and time.
``Train'' for example, links to a parent node, ``Rail Transport,''
while many child nodes such as ``Steel'' and ``Horsepower'' link to ``Train.''
As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
contains ``Goods,'' ``Economics,'' ``Social Science,''
leading ultimately to ``Philosophy''.
Unlike previous taxonomies created by individuals
\cite{locke, descartes, aristotle, hist_thesaurus},
the relations in the FLN emerge without a centralized effort 
as the aggregate of each article's authors choice of first link.

@@

\section{Traversing the First Link Network}

An essential feature of a directed network's structure is the degree distribution 
\cite{newman}. 
The degree distribution has been used to study many phenomena from disease outbreak 
\cite{disease} 
to the dynamics of social networks 
\cite{social_nets}.
The in-degree distribution in the FLN describes how many first links point to a 
particular article. 
Articles with zero in-degree have no references---they are outer leaves in the FLN 
or are disconnected entirely.

@@

Starting at each article, we construct a path through the FLN, 
to map the flow of connections among articles.
The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
Previous studies have used flow to characterize the structure of river networks
\cite{geo_basins, dodds} and describe the organization of 
blood networks, food systems, and transportation networks
\cite{food_webs}.

@@

We can characterize the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
(Fig.~\ref{fig:Traversal Visits}), the path starting at article A is the 
first column in the traversal visits matrix. 
An entry of $1$ indicates the path contains a given article and 
$0$ indicates the path does not.
To compute the number of traversal visits for an article, we sum the corresponding row in our matrix.
The traversal visits matrix for 
our Wikipedia dataset consists of 121 million entries encoding each path out of the more than 
googol possible paths ($4.7*10^6*2^{4.7*10^6}$) through the FLN .

By measuring the number of first links between
two articles, we obtain an additional piece of information we call 
{\it path length.} We can compute the length of a path by summing 
along a column in our traversal visits matrix. 
In our Wikipedia dataset, the sum of all path lengths is 232 million first links.
The path length describes how closely related topics are.
Although ``Train'' is related to ``Economics'' for example, there are several articles
bridging the connection:
``Train'' is more specifically related to transportation, whose object
is often goods. Goods are one of the fundamental objects of study in Economics.
Described in links, this relationship is captured by a relatedness of $4$ 
first links ultimately connecting ``Train'' to ``Economics.''

@@

While traversal visits measure accumulation, each article's first link also 
influences the shape of the FLN. 
At a point of accumulation, a single article's first link 
can exert great influence over the shape of the FLN by directing many
references on a particular path. To distinguish between an article 
that simply happened to fall within a cycle from an article funneling 
many first links, we develop a second metric called {\it traversal funnels}.

To measure traversal funnels, we traverse the FLN in the same manner as we 
did for traversal visits, but end a path once we enter a cycle.
We are then able to distinguish between an article related to many other ideas
only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
An article with a large number of traversal funnels directs many references
into a particular cycle. In our sample network 
(see Fig.~\ref{fig:Traversal Funnels}) article C 
directs the flow of links towards the 3-cycle, while articles A and B are 
recipients of the flow---without exerting direct influence themselves.

By studying the FLN not only as collection of directly linked pairs of articles, but
as a flow, 
we build a powerful arsenal of information.
From accumulated references, cycles, and basins, we can measure how the many articles in Wikipedia are organized and related.

@@

We rank all 11 million articles by in-degree to find 
the ``United States'' with 80,249 direct first links as the most referenced
Wikipedia article 
(Fig.~\ref{fig:indegree list}). 
Other high-ranking articles
include foundational abstract concepts such as ``village,'' ``species,''; 
sports associations such as ``American Football,'' ``Association Football''; 
and developed nations such as ``France,'' ``Japan,'' ``Russia,'' ``Australia,'' and 
the ``Netherlands.'' These high ranking articles are useful abstractions: nations
describe a collection of individuals with a common culture, language, or 
geographical proximity; sports teams describe an ever changing collection of 
players often associated with a cultural identity or a geographical 
region. 
Since abstractions such as nations and teams are inherently comprised
of many parts, authors reference the abstraction when describing a part.
In describing the New England Patriots or the New York Giants
for example, ``American Football'' is the natural abstraction, which 
anchors many specific teams.

``Philosophy'' and other philosophical concepts
are not among the highest-ranking articles by in-degree.
``Philosophy'' has an in-degree of only 581, with direct first links from articles about Philosophers and areas of Philosophy: ``Existentialism and Humanism,'' ``Predeterminism,'' ''Synoptic Philosophy,'' ``Qualia,'' ``Dorothy Emmet,'' and ``Christopher W. Morris.''
While many articles accumulate at ``Philosophy'' (see traversal visits discussion below), 
the accumulation is not the 
result of many articles directly referencing ``Philosophy.'' 
Instead, first links flow towards ``Philosophy'' as the 
ultimate anchor, by generalizing from specific to broad.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
  \caption{
    \textbf{Highest ranking articles by number of traversal visits.}
We compute the number of traversal visits for each article in the FLN (see 
Traversal Algorithm section for details). In doing so, we can rank each article
by the accumulation of first links.
The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
  \label{fig:highest visits}
\end{figure}

This roughly corresponds to the official number of articles on Wikipedia: 
$~4.7$ million as of November 2014---approximately half of the 11 million 
articles in the XML dump are redirects or disambiguations, not full articles.
The most common path length is 29, with an interquartile range of 4---path length
is typically between 26 and 29 articles.
As a distribution, more than $75\%$ of articles have a path length below 
$50$ first links 
while a few temporally organized paths exceed 50 links 
(see Fig.~\ref{fig:Path Length Distribution}).

@@

\begin{figure*}[tph!]
  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
  \caption{
    \textbf{Highest ranking 2-cycles and 3-cycles.}
We identify pairs of articles whose first links point to one another, forming
a 2-cycle. We then rank each pair of articles by the total number of 
traversal visits to gauge the most referenced groups of two articles linked
to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
same concepts as opposed to distinct concepts. Similarly, we identify and 
rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
  \label{fig:2-cycles}
\end{figure*}
In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
to two regimes corresponding to power law rank exponents of 
$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
On the aggregate, the distribution suggests
a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' `` Geography,'' and ``Philosophy''
as well as ``Community,'' ``State,'' ``Earth,'' ``Information,'' ``Power,'' and ``Communication.''
The first links flow towards broader topics, accumulating at of these foundational notions.
While ``Banana'' is a concrete fruit for example, 
the first links flow from ``Fruit'' to ``Botany'' to ``Biology,'' and ultimately 
culminate with ``Science'' and ``Philosophy.'' 
The first links anchor the wealth of specific knowledge on Wikipedia to 
a few notions foundational to society.

@@

We identify 2-cycles, meaning a pair of articles with first link pointing to one another.
Of the 11 million articles, roughly 84,000 are memebers of 2-cycles. 
The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
``Health Care'' and ``Medical Case Management,'' ``Broadcasting House,'' and ``BBC,'' ``Secondary Education'' and ``Secondary School'' 
(see Fig.~\ref{fig:2-cycles}).

Outside of the highest ranking 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related concepts. 
We also observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').

Similarly, 3-cycles capture a synonymous or close relation among 3 articles: ``Tree of life (Biology),'' ``Tree of life (disambiguation),'' 
and ``Tree of life''; ``Cinema of India,'' ``Indian Cinema,'' and ``Telugu Cinema''
(see Fig.~\ref{fig:3-cycles}).
Once we extend our cycle size beyond a length of 6 however, 
``Philosophy'' along with the remaining list of high ranking articles by traversal visits dominate.
The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
Other lengthy cycles span $60-75$ articles including collections of articles on national histories such as ``Japanese Eras'' 
or judicial bodies such as the ``Legislative Assembly of Ontario.''

@@

We can group articles lying on the same path to identify {\it basins}: 
a group of path-connected articles.
Since cycles identify only groups of articles with a closed set of links, 
we additionally measure and rank basins to capture groups of closely related
articles branching outside of a cycle into the rest of the FLN.
We rank basins by the total number of traversal visits for each article in the path. 
Akin to river networks, these basins are areas of accumulation with a path 
flowing outwards to the rest of the FLN.

The highest ranking basins by the number of traversal visits are groups of articles
around ``Philosophy.'' 
The highest ranking paths include branches of philosophy flowing through 
``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy.'' Other paths
include concepts around ``Mathematics,'' ``Scientific,'' ``Experiments,'' 
``Biology,'' and ``Fact.''
These paths link many specific articles to ``Philosophy,'' each funneled through a particular domain.

Moving beyon basins around ``Philosophy,'' we find other basins around 
foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
``Presentation,'' and ``Belief System.'' 
The basins around each of these foundational notions are 
various paths containing related articles. For example around 
``Community'' we find basins
flowing from ``United States'' to ``Federal Republic'' to ``Political Union'' to ``State'' culminating at ``Community''; we also find basins flowing from 
``Public Policy'' to ``Executive (government)'' to ``Government'' to ``State'' and then 
to ``Community''; we also find paths flowing through a similar chain beginning
with ``Democracy,'' another beginning with ``Constitution,'' another at 
``Dictatorship,'' and so on. The articles build from specific means of organizing
a community (or society) and then build up to ``Community.'' 
Other basins around landmass for example begin at specific geographical regions
such as ``Eastern Europe'' building up to ``Continent'' and finally ``Landmass''.

@@

\subsection{Traversal Funnels}

To analyze the influence an article exerts in shaping the 
structure of the FLN, we compute the number of traversal funnels for each.
Articles directing more paths exert a greater influence over the structure
of the FLN by increasing the accumulation of first links
on a particular path. By measuring traversal funnels, we distinguish between an article that simply happened to fall within a cycle from an article funneling 
many first links.

Ranking articles by the number of traversal funnels we find 
``Philosophy'' to be by far the highest-ranking article with 
$7.37$ million paths
(see Fig.~\ref{fig:Funnels}).
Of any article, the number of traversal funnels Philosophy holds exceeds 
all others by more than two orders of magnitude.
The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
and similar articles accumulates the overwhelming proportion of its 
references through ``Philosophy'': $7.37$ million of the $7.4$ million references
are funneled through ``Philosophy''.
Second on the list of highest-ranking articles by traversal funnels is 
``Presentation'' with only $30$ thousand paths. Similarly abstract 
concepts also rank highly such as ``Tree of life'' (30 thousand), 
``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).

@@

@@

\end{figure}
As a distribution, we find few articles influence the structure of the 
FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
more than $99\%$ with none---most articles are recipients of 
the references flowing through the articles with at least one traversal funnel.
When fit to a log-log linear model we find the $99\%$ of articles with zero
traversal funnels form one regime (with log(rank) less than 4).
The top regime, corresponding to the $17, 821$ articles with at least one 
traversal funnel strongly fits a linear model with
a Pearson's Correlation Coefficient of $-0.99$ and a 
corresponding power law rank exponent $\alpha = -1.08$
and a size exponent $\gamma = 0.074$ 
(see Fig.~\ref{fig:Funnels Distribution}). Even within the few articles
influencing the structure of the FLN, only a handful of these exert most of the 
control.

@@

\section{Appendix}
\label{Appendix}

\subsection{Constructing The First Link Network}

>>>>

@@

Wikipedia is a towering achievement of the Internet age. 
At no point in history has a larger or more meticulously indexed collection of human knowledge 
existed.
Wikipedia contains 37 million articles in 283 languages, 
with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
\cite{clauson2008scope, stats}.
Demonstrating its relevance to modern inquiry,
Wikipedia is the sixth most visited
site in the world, surpassing 18 billion page views and 10 million edits in January, 2013 alone.
\cite{wiki_edits, wiki_views}.

Wikipedia has naturally become the object of many studies. 
Researchers have examined the cultural dynamics among editors
\cite{iba2010analyzing, samoilenko2016linguistic},
the accuracy of the content relative to traditional encyclopedias
\cite{holman2008comparison, giles2005internet},
the topics covered 
\cite{halavais2008analysis},
and bias against portions of the population
\cite{hill2013wikipedia}.
Wikipedia's content has also proven to be a powerful tool. 
Researchers have used Wikipedia to identify missing dictionary entries
\cite{williams2015identifying},
cluster short text
\cite{banerjee2007clustering},
compute semantic relatedness
\cite{gabrilovich2007computing},
and disambiguate meaning \cite{cucerzan2007large}.

While these many studies have dissected and fruitfully applied Wikipedia's content,
here we examine the connections among the many articles.
A hyperlink from one Wikipedia article to another
naturally indicates a relationship between the two articles
\cite{kamps2009wikipedia}.
The notion that hyperlinks convey information about the content of a
page has proved enormously successful in multiple domains from search engine algorithms 
such as PageRank 
\cite{page1999pagerank} 
to topic classification
\cite{chakrabarti2001integrating}.
Here, we treat a hyperlink as a mechanism connecting two topics.

@@

By following the first hyperlink contained in all articles of the English edition of
Wikipedia as of November, 2014, we form a directed network:
{\it Wikipedia's First Link Network} (FLN).
Inspired by the claim that the majority of first links lead to
``Philosophy''---popularized by an 
\href{https://xkcd.com/903/}{xkcd}
comic and subsequently
discussed in blog posts 
\cite{xkcd, mat_blog, Ilmari_first_links, 
bob_west}
---
we holistically study how the FLN  yields a flow of connections.
For methodological details, see Sec.~\ref{Appendix}.

The FLN is a wealth of relations among inventions, places,
figures, objects, and events across space and time.
``Train'' for example, links to a parent node, ``Rail Transport,''
while many child nodes such as ``Steel'' and ``Horsepower'' link to ``Train.''
As shown in Fig.~\ref{fig:Train First Links}, the path starting at ``Train'' 
contains ``Goods,'' ``Economics,'' ``Social Science,''
leading ultimately to ``Philosophy''.
Unlike previous taxonomies created by individuals
\cite{Bolton2007, descartes, aristotle, hist_thesaurus},
the relations in the FLN emerge without a centralized effort 
as the aggregate of each article's authors choice of first link.

@@

\section{Traversing the First Link Network}
\label{Traversal Algorithm}

An essential feature of a directed network's structure is the degree distribution 
\cite{newman2003structure}. 
The degree distribution has been used to study many phenomena from disease outbreak 
\cite{eubank2004modelling} 
to the dynamics of social networks 
\cite{newman2002random}.
The in-degree distribution in the FLN describes how many first links point to a 
particular article. 
Articles with zero in-degree have no references---they are outer leaves in the FLN 
or are disconnected entirely.

@@

Starting at each article, we construct a path through the FLN, 
to map the flow of connections among articles.
The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
Previous studies have used flow to characterize the structure of river networks
\cite{horton1945erosional, dodds1999unified} and describe the organization of 
blood networks, food systems, and transportation networks
\cite{garlaschelli2003universal}.

@@

We can characterize the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
(Fig.~\ref{fig:Traversal Visits}), the path starting at article A is the 
first column in the traversal visits matrix. 
An entry of $1$ indicates the path contains a given article and 
$0$ indicates the path does not.
To compute the number of traversal visits for an article, we sum the corresponding row in our matrix.
The traversal visits matrix for 
our Wikipedia dataset consists of 121 million entries encoding each path out of the more than 
googol possible paths ($\simeq4.7\times10^6\times2^{4.7\times10^6}$) through the FLN .

By measuring the number of first links between
two articles, we obtain an additional piece of information we call 
{\it path length}, computed by summing 
columns in the traversal visits matrix. 
Path length describes how closely related topics are.
Although ``Train'' is related to ``Economics'' for example, there are several articles
bridging the connection:
``Train'' is more specifically related to transportation, whose object
is often goods. Goods are one of the fundamental objects of study in Economics.
Described in links, this relationship is captured by a relatedness of $4$ 
first links ultimately connecting ``Train'' to ``Economics.''

@@

While traversal visits measure accumulation, each article's first link also 
influences the shape of the FLN. 
At a point of accumulation, a single article's first link 
can exert great influence over the shape of the FLN by directing many
references on a particular path. To distinguish between an article 
that simply happened to fall within a cycle from an article funneling 
many first links, we develop a second metric called {\it traversal funnels}.
To measure traversal funnels, we traverse the FLN in the same manner as we 
did for traversal visits, but end a path once we enter a cycle.
We are then able to distinguish between an article related to many other ideas
only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
In our simple network 
(see Fig.~\ref{fig:Traversal Funnels}), article C 
directs the flow of links towards the 3-cycle, while articles A and B are 
recipients of the flow---without exerting direct influence themselves.

By considering the FLN not only as collection of directly linked pairs of articles, but
as a flow, 
With accumulated references, cycles, and basins, we can explore how the many articles in Wikipedia are organized and related.

@@

We rank all 11 million articles in the FLN by in-degree to find 
the ``United States'' with 80,249 direct first links as the most referenced
Wikipedia article 
(Fig.~\ref{fig:indegree list}). 
Other high-ranking articles
include foundational abstract concepts such as ``village,'' ``species,''; 
sports associations such as ``American Football,'' ``Association Football''; 
and developed nations such as ``France,'' ``Japan,'' ``Russia,'' ``Australia,'' and 
the ``Netherlands.'' These high ranking articles are useful abstractions: nations
describe a collection of individuals with a common culture, language, or 
geographical proximity; sports teams describe an ever changing collection of 
players often associated with a cultural identity or a geographical 
region. 
Since abstractions such as nations and teams are inherently comprised
of many parts, authors often reference the abstraction when describing a part.

``Philosophy'' and other philosophical concepts
are not among the highest-ranking articles by in-degree.
``Philosophy'' has an in-degree of only 581, with direct first links from articles about Philosophers and areas of Philosophy: ``Existentialism and Humanism,'' ``Predeterminism,'' ''Synoptic Philosophy,'' ``Qualia,'' ``Dorothy Emmet,'' and ``Christopher W. Morris.''
So while a great many FLN paths flow to 
``Philosophy'' (see traversal visits discussion below), 
the accumulation is not the 
result of many articles directly referencing ``Philosophy.'' 
Instead, first links flow towards ``Philosophy'' as the 
ultimate anchor, by generalizing from specific to broad.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
  \caption{
    \textbf{Highest ranking articles by number of traversal visits.}
We compute the number of traversal visits for each article in the FLN (see 
Sec.~\ref{Traversal Algorithm}). 
In doing so, we can rank each article by the accumulation of first links.
The highest ranking articles by traversal visits reveal where the greatest accumulation occurs.}
  \label{fig:highest visits}
\end{figure}

This roughly corresponds to the official number of articles on Wikipedia: 
$~4.7$ million as of November 2014---approximately half of the 11 million 
articles in the XML dump are redirects or disambiguations, not full articles.
The most common path length is 29, with $75\%$ of path lengths ranging between 7--30 articles.
As a distribution, more than $75\%$ of articles have a path length below 
$50$ first links 
while a few temporally organized paths exceed 50 links 
(see Fig.~\ref{fig:Path Length Distribution}).

@@

\begin{figure*}[tph!]
  \includegraphics[width=\textwidth]{graphics/2_and_3_cycles.pdf}
  \caption{
    \textbf{Highest ranking 2-cycles and 3-cycles.}
We identify pairs of articles whose first links point to one another, forming
a 2-cycle. We then rank each pair of articles by the total number of 
traversal visits to gauge the most referenced groups of two articles linked
to each other. We find 2-cycles often capture synonyms or articles representing nearly the 
same concepts as opposed to distinct concepts. Similarly, we identify and 
rank 3-cycles to find they appear to capture three closely related ideas or synonyms.}
  \label{fig:cycles}
\end{figure*}
In Fig.~\ref{fig:Distribution of Visits}, we fit traversal visits against rank
to two regimes corresponding to power law rank exponents of 
$\alpha = -1.23 (\gamma = 0.187$) and $\alpha = -0.579 (\gamma = -0.727)$.
On the aggregate, the distribution suggests
a handful of the highest ranking articles contain a disproportionate number of traversal visits, while most have none. The skew in the distribution is not terribly surprising when considering the heuristic of how the links flow: from specific to general.

The highest ranking articles by traversal visits are broad topics spanning academic disciplines 
or notions fundamental to society~(Fig. \ref{fig:highest visits}): ``Science,'' ``Mathematics,'' ``Geography,'' and ``Philosophy''
as well as ``Community,'' ``State,'' ``Earth,'' ``Information,'' ``Power,'' and ``Communication.''
The first links flow towards broader topics, accumulating at of these foundational notions.
While ``Banana'' is a concrete fruit for example, 
the first links flow from ``Fruit'' to ``Botany'' to ``Biology,'' and ultimately 
culminate with ``Science'' and ``Philosophy.'' 
We observe that first links direct the wealth of specific knowledge on Wikipedia to 
a few foundational notions.

@@

We first identify 2-cycles, meaning a pair of articles with first link pointing to one another.
Of the 11 million articles, roughly 84,000 are members of 2-cycles. 
The highest ranking 2-cycles by traversal visits tend to be synonyms (or nearly so) rather than distinct, yet connected topics:
``Health Care'' and ``Medical Case Management''; ``Broadcasting House'' and ``BBC''; and ``Secondary Education'' and ``Secondary School'' 
(see Fig.~\ref{fig:cycles}).

Outside of these prominent 2-cycles, the typical 2-cycle signals a connection between distinct, yet very closely related concepts. 
For example, we observe link patterns such as inventor to product (``Voere'' to ``VEC-91''), event to organizer (``Poetry Bus Tour'' to ``Weave Books''), and book to author (``Anatomy of Britain'' to ``Anthony Sampson'').

Similarly, 3-cycles capture a synonymous or close relation among 3 articles: ``Tree of life (Biology),'' ``Tree of life (disambiguation),'' 
and ``Tree of life''; ``Cinema of India,'' ``Indian Cinema,'' and ``Telugu Cinema''
(see Fig.~\ref{fig:cycles}).
Once we extend our cycle size beyond a length of 6 however, 
``Philosophy'' along with the remaining list of high ranking articles by traversal visits dominate.
The longest cycle in the network spans 365 articles of Eastern Orthodox Liturgics for each calendar day.
Other lengthy cycles span 60--75 articles including collections of articles on national histories such as ``Japanese Eras'' 
or judicial bodies such as the ``Legislative Assembly of Ontario.''

@@

The highest ranking basins (Sec.~\ref{Traversal Algorithm}) by the number of traversal visits are groups of articles around ``Philosophy.'' 
The highest ranking paths include branches of philosophy flowing through 
``Awareness,'' ``Existence,'' and ``Consciousness'' to ``Philosophy,'' 
which form a cycle of seven articles. As of April 10, 2016 the basin flowing from 
``Philosophy'' has changed to contain instead ``Ancient Greece,'' ``Greek,'' and 
``Indo-European Languages'' (see network resilience Sec.~\ref{Traversal Funnels}).
Other paths
include concepts around ``Mathematics,'' ``Scientific,'' ``Experiments,'' 
``Biology,'' and ``Fact.''
These paths link many specific articles to ``Philosophy,'' each funneled through a particular domain.

Moving beyond basins around ``Philosophy,'' we find other basins around 
foundational concepts such as ``Community,'' ``Landmass,'' ``Federal Government,'' 
``Presentation,'' and ``Belief System.'' 
The basins around each of these foundational notions are 
various paths containing related articles. For example around 
``Community'' we find basins
flowing from ``United States'' to ``Federal Republic'' to ``Political Union'' to ``State'' culminating at ``Community''; we also find basins flowing from 
``Public Policy'' to ``Executive (government)'' to ``Government'' to ``State'' and then 
to ``Community''; we also find paths flowing through a similar chain beginning
with ``Democracy,'' another beginning with ``Constitution,'' another at 
``Dictatorship,'' and so on. The articles build from specific means of organizing
a community (or society) and then build up to ``Community.'' 
Other basins around landmass for example begin at specific geographical regions
such as ``Eastern Europe'' building up to ``Continent'' and finally ``Landmass''.

@@

\subsection{Traversal Funnels}
\label{Traversal Funnels}

To analyze the influence an article exerts in shaping the 
structure of the FLN, we compute the number of traversal funnels for each.
Ranking articles by the number of traversal funnels we find 
``Philosophy'' to be by far the highest ranking article with 
$7.37$ million paths
(see Fig.~\ref{fig:Funnels}).
Of any article, the number of traversal funnels Philosophy holds exceeds 
all others by more than two orders of magnitude.
The ``Philosophy'' cycle which contains ``Existence,'' ``Awareness,'' ``Reality,'' 
and similar articles accumulates the overwhelming proportion of its 
references through ``Philosophy'': $7.37$ million of the $7.4$ million references
are funneled through ``Philosophy''.
Second on the list of highest-ranking articles by traversal funnels is 
``Presentation'' with only $30$ thousand paths. Similarly abstract 
concepts also rank highly such as ``Tree of life'' (30 thousand), 
``Reality'' (13 thousand), and ``Jurisdiction'' (3 thousand).

@@

Traversal funnels are comparatively resilient to changes in the FLN over time. While a few links can alter a cycle, potentially changing the number of traversal visits for an article, traversal funnels do not depend on cycles. Even when the particular links between articles change, the overall influence measured by traversal funnels is relatively resilient.

@@

\end{figure}
As a distribution, we find few articles influence the structure of the 
FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
more than $99\%$ with none---most articles are recipients of 
the references flowing through the articles with at least one traversal funnel.
When fit to a log-log linear model we find the $99\%$ of articles with zero
traversal funnels form one regime (with $\log_{10}$(rank) less than 4).
The top regime, corresponding to the $17, 821$ articles with at least one 
traversal funnel strongly fits a linear model with
a Pearson's Correlation Coefficient of $-0.99$ and a 
corresponding power law rank exponent $\alpha = -1.08$
and a size exponent $\gamma = 0.074$ 
(see Fig.~\ref{fig:Funnels Distribution}). Even within the few articles
influencing the structure of the FLN, only a handful of these exert most of the 
control.

@@
